---
title: "Used Cars Cleaning Notebook"
output: html_notebook
---

This is the cleaning notebook for the used cars dataset. The dataset comes from kaggle: ('https://www.kaggle.com/datasets/austinreese/craigslist-carstrucks-data')

Once the dataset is thoroughly cleaned, an R Shiny dashboard will be created from the data.

```{r, results = "hide", message = FALSE, warning = FALSE}
### load tidyverse library
library(tidyverse)

### read in data
df <- read_csv("vehicles.csv")
```

```{r, message = FALSE, warning = FALSE}
### Check shape of the dataset
df %>% dim()
```

The dataset has been sucessfully loaded and contains 426k rows and 26 columns of data

## Cleaning
- Explore the dataset and remove unwanted columns and Null values

```{r, message = FALSE, warning = FALSE}
### Examine columns
df %>% colnames()
```

```{r, message = FALSE, warning = FALSE}
### Examine the first 5 rows that do not have any nulls in the column 'year'
df %>%
  filter(!is.na(year)) %>%
  head(5)
```
Rows that we don't need
- url, region_url, VIN, image_url, description, lat, long, posting_date
Look into
- drive, size, type

```{r, message = FALSE, warning = FALSE}
### Examine the drive column and count
df <- df %>%
  select(-url, -region_url, -VIN, -image_url, -description, -lat, -long, -posting_date)

```


```{r, message = FALSE, warning = FALSE}
### Examine the drive column and count
df %>%
  count(drive) %>%
  mutate(proportion = n / sum(n))

```

About 30% of these rows do not have a designated driving type, that is a large chunk of the data set, lets explore more and cut down on others and come back to this column and see what is left

```{r, message = FALSE, warning = FALSE}
### Examine the drive column and count
df %>%
  count(size) %>%
  mutate(proportion = n / sum(n))

```
70% of the data does not have a designated size
- we might just remove this column?

```{r, message = FALSE, warning = FALSE}
### Examine the drive column and count
df %>%
  count(type) %>%
  mutate(proportion = n / sum(n))

```

There are no Nulls in the type column, but most cars are either sedans and SUVs
- We will probably get rid of non-normal vehicles like bus/offroad

```{r, message = FALSE, warning = FALSE}
### Examine columns
df %>% colnames()
```


I want to keep all of these columns if we can reliably fill the missing data


```{r, message = FALSE, warning = FALSE}
### Examine columns
df %>%
  summarise_all(~ sum(is.na(.))) %>%
  pivot_longer(cols = everything(), names_to = "Column", values_to = "NA_Count") %>%
  mutate(Proportion = NA_Count / nrow(df))
```

### Year Column
 
```{r, message = FALSE, warning = FALSE}
### Examine columns
df['year'] %>% summary()
```



```{r, message = FALSE, warning = FALSE}
### Examine distribution of year column
df %>% 
  count(year) %>%
  mutate(proportion = n / sum(n)) %>%
  ggplot(aes(x = year, y = n)) + 
  geom_col(fill = "#0073C2FF", alpha = 0.8, width = 0.7) + 
  geom_vline(xintercept = 2000, color = "red", linetype = "dashed", size = 1) +
  labs(
    title = "Distribution of Vehicles by Year",
    x = "Year",
    y = "Count"
  ) +
  theme_minimal(base_size = 15)
```

Remove data from before 2000?
- most cars being sold are in the range of 2008-2017 


```{r, message = FALSE, warning = FALSE}
### Examine distribution of year column
df <- df %>% 
  filter(year >= 2000)
```

```{r, message = FALSE, warning = FALSE}
### Examine distribution of year column
df %>% 
  count(year) %>%
  mutate(proportion = n / sum(n)) %>%
  ggplot(aes(x = year, y = n)) + 
  geom_col(fill = "#0073C2FF", alpha = 0.8, width = 0.7)+
  labs(
    title = "Distribution of Vehicles by Year",
    x = "Year",
    y = "Count"
  ) +
  theme_minimal(base_size = 15)
```

### Price Column
Clean Price Column


```{r, message = FALSE, warning = FALSE}
### Examine distribution of the  price column
df %>%
  count(price) %>%
  mutate(price_bucket = cut(
    price,
    breaks = c(-Inf, seq(2000, 80000, by = 2000), Inf),
    labels = c("<=2k", paste(seq(2, 78, by = 2), "k-", seq(4, 80, by = 2), "k", sep = ""), ">80k"),
    right = TRUE
  )) %>%
  ggplot(aes(x = price_bucket)) + 
  geom_bar(fill = "#0073C2FF", color = "black", alpha = 0.7) + 
  labs(
    title = "Histogram of Vehicle Prices",
    x = "Price Bucket",
    y = "Count"
  ) + 
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text = element_text(color = "black"),
    axis.title = element_text(face = "bold")
  ) +
  scale_x_discrete(drop = FALSE)
```

```{r, message = FALSE, warning = FALSE}
### Examine distribution of year column
df %>%
  mutate(price_bucket = cut(
    price,
    breaks = c(-Inf, seq(2000, 80000, by = 2000), Inf),
    labels = c("<=2k", paste(seq(2, 78, by = 2), "k-", seq(4, 80, by = 2), "k", sep = ""), ">80k"),
    right = TRUE
  )) %>%
  count(price_bucket) %>%
  mutate(proportion = n / sum(n)*100)
  
```

about 12.8% of the data contains observations where the price of the car is less than 2000 dollars
- about 3.4% have prices listed >80k, i am going to remove these

```{r, message = FALSE, warning = FALSE}
### Examine distribution of year column
df <- df %>% 
  filter(price < 80000)
```


check the price column again
- use increments of 500

```{r, message = FALSE, warning = FALSE}
### Examine distribution of the  price column
df %>%
  count(price) %>%
  mutate(price_bucket = cut(
    price,
    breaks = c(-Inf, seq(500, 80000, by = 500), Inf),
    labels = c("<=500", paste(seq(0.5, 79.5, by = 0.5), "k-", seq(1, 80, by = 0.5), "k", sep = ""), ">80k"),
    right = TRUE
  )) %>%
  ggplot(aes(x = price_bucket)) + 
  geom_bar(fill = "#0073C2FF", color = "black", alpha = 0.7) + 
  labs(
    title = "Histogram of Vehicle Prices (500 Increments)",
    x = "Price Bucket",
    y = "Count"
  ) + 
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text = element_text(color = "black"),
    axis.title = element_text(face = "bold")
  ) +
  scale_x_discrete(drop = FALSE)
```


```{r, message = FALSE, warning = FALSE}
### Examine distribution of year column
df %>%
  mutate(price_bucket = cut(
    price,
    breaks = c(-Inf, seq(500, 80000, by = 500), Inf),
    labels = c("<=500", paste(seq(0.5, 79.5, by = 0.5), "k-", seq(1, 80, by = 0.5), "k", sep = ""), ">80k"),
    right = TRUE
  )) %>%
  count(price_bucket) %>%
  mutate(proportion = n / sum(n)*100)
  
```
19% of the cars have prices less than 1000$, i will get rid of these

```{r, message = FALSE, warning = FALSE}
### Examine distribution of year column
df <- df %>% 
  filter(price > 1000)
```


```{r, message = FALSE, warning = FALSE}
### Examine distribution of the  price column
df %>%
  count(price) %>%
  mutate(price_bucket = cut(
    price,
    breaks = c(-Inf, seq(500, 80000, by = 500), Inf),
    labels = c("<=500", paste(seq(0.5, 79.5, by = 0.5), "k-", seq(1, 80, by = 0.5), "k", sep = ""), ">80k"),
    right = TRUE
  )) %>%
  ggplot(aes(x = price_bucket)) + 
  geom_bar(fill = "#0073C2FF", color = "black", alpha = 0.7) + 
  labs(
    title = "Histogram of Vehicle Prices (500 Increments)",
    x = "Price Bucket",
    y = "Count"
  ) + 
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text = element_text(color = "black"),
    axis.title = element_text(face = "bold")
  ) +
  scale_x_discrete(drop = FALSE)
```

ok this looks good to me, i guess...


### Region Column

```{r, message = FALSE, warning = FALSE}
### Examine the drive column and count
df %>%
  count(region) %>%
  mutate(proportion = n / sum(n))

```

There are too many regions, they span different cities and subdivisions in the USA
- if i were to keep this column it would be too many variables to account for.
- i am going to drop this column

```{r, message = FALSE, warning = FALSE}
### Examine the drive column and count
df %>%
  select(-region)

```


### manufacturer Column

```{r, message = FALSE, warning = FALSE}
### Examine the drive column and count
df %>%
  count(manufacturer) %>%
  mutate(proportion = n / sum(n))

```

This is an important column to keep, i need to remove cars without a manufacturer
- and also remove manufacturer in which we don't have many observations with (<1000)
- aston-martin 16
- ferrari 3
- harley-davidson 124
- land rover 9
- morgan 1
- mercury 773
- tesla 799
- saturn 984
- NA


